Exploring how changes in cause of death have narrowed the black-white life expectancy gap in the United States.

### 2017-02-22 Code/combine_results.Rmd

---
title: "Combine_results"
author: "Corinne Riddell"
date: "2/9/2017"
output: html_document
---

@@

completed.states <- included.states[17:24]
```

@@

for(state_i in 1:length(completed.states)){

@@

COD.levels <- c("Cardiovascular", "Cancers", "Communicable", 
                "Non-communicable", "Injuries", "All other causes")

cod_decomp_results$COD <- factor(cod_decomp_results$COD, COD.levels)
age_cod_results$COD <- factor(age_cod_results$COD, levels = 1:6, 
                               labels = c("All other causes", "Cancers", "Cardiovascular", "Communicable",
                                          "Injuries", "Non-communicable"))

@@

```{r save_combined_datasets, echo = F}
write.csv(age_cod_results, "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_results.csv")
write.csv(age_decomp_results, "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_decomp_results.csv")
write.csv(cod_decomp_results, "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/cod_decomp_results.csv")
write.csv(BlackWhite_results, "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/BlackWhite_results.csv")

>>>>

---
title: "Combine_results"
date: "2/9/2017"
output: html_document
---

@@

completed.states <- included.states[c(1:4, 17:24)]
```

@@

for(state_i in 2:length(completed.states)){

@@

# COD.levels <- c("Cardiovascular", "Cancers", "Communicable", 
#                 "Non-communicable", "Injuries", "All other causes")

#cod_decomp_results$COD <- factor(cod_decomp_results$COD, COD.levels)
age_cod_results$COD <- factor(age_cod_results$COD, levels = 1:6, 
                               labels = c("All other causes", "Cancers", "Cardiovascular", "Communicable",
                                          "Injuries", "Non-communicable"))

@@

```{r save_combined_datasets, echo = F}
write.csv(age_cod_results, "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_results.csv")
write.csv(age_decomp_results, "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_decomp_results.csv")
write.csv(cod_decomp_results, "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/cod_decomp_results.csv")
write.csv(BlackWhite_results, "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/BlackWhite_results.csv")
write.csv(cod_change_results, "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/cod_change_results.csv")

### 2017-02-22 Manuscript/Manuscript_COD_decomp.Rmd

```{r read_data, echo=F}
age_cod_results <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_results.csv")
age_decomp_results <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_decomp_results.csv")
cod_decomp_results <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/cod_decomp_results.csv")
BlackWhite_results <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/BlackWhite_results.csv")

COD.levels <- c("Cardiovascular", "Cancers", "Communicable", 
                "Non-communicable", "Injuries", "All other causes")

cod_decomp_results$COD <- factor(cod_decomp_results$COD, COD.levels)
age_cod_results$COD <- factor(age_cod_results$COD, levels(age_cod_results$COD)[c(3, 2, 4, 5, 6, 1)])

age_decomp_results$age <- factor(age_decomp_results$age, levels = levels(age_decomp_results$age)[c(1, 2, 11, 3:10, 12:19)])
age_cod_results$age <- factor(age_cod_results$age, levels = levels(age_cod_results$age)[c(1, 2, 11, 3:10, 12:19)])                                                             
```

@@

bw_dat <- BlackWhite_results %>% 
  filter(year %in% c(year1, year2, year3, year4), is.na(LE_wbgap) == F) %>%
  select(state, year, LE_wbgap, sex) %>%
  arrange(year)

bw_dat <- tidyr::spread(bw_dat, year, LE_wbgap) 
names(bw_dat)[3:6] <- c("first.gap","second.gap", "third.gap", "fourth.gap")  
bw_dat.males <- bw_dat %>% filter(sex == "Male") %>%
  mutate(era1.diff = first.gap - second.gap,
         era2.diff = second.gap - third.gap,
         era3.diff = third.gap - fourth.gap,
         state.gdiff.order.1 = reorder(state, first.gap),
         state.gdiff.num.1 = as.numeric(state.gdiff.order.1),
         state.gdiff.order.2 = reorder(state, second.gap),
         state.gdiff.num.2 = as.numeric(state.gdiff.order.2),
         state.gdiff.order.3 = reorder(state, third.gap),
         state.gdiff.num.3 = as.numeric(state.gdiff.order.3)
  )

@@

cod.df <- data.frame(state = factor(), sex = factor(), era = character(), 
                     COD = factor(), Contribution.to.change = numeric(), 
                     Contrib.to.change.prop = numeric(), narrowed_gap = logical())

age.df <- data.frame(state = factor(), sex = factor(), era = character(), 
                     age_minbin = numeric(), Contribution.to.change = numeric(), 
                     Contrib.to.change.prop = numeric(), narrowed_gap = logical())

years <- c(1969, 1983, 1993, 2013)

for (state.i in levels(BlackWhite_results$state)) {
  for (sex.i in c("Male", "Female")) {
    for (year.i in 1:3) {

      table1 <- cod_decomp_results %>% filter(state == state.i & sex == sex.i & year == years[year.i]) %>% arrange(COD)
      table2 <- cod_decomp_results %>% filter(state == state.i & sex == sex.i & year == years[year.i + 1]) %>% arrange(COD)

      table3 <- age_decomp_results %>% filter(state == state.i & sex == sex.i & year == years[year.i]) %>% arrange(age_minbin)
      table4 <- age_decomp_results %>% filter(state == state.i & sex == sex.i & year == years[year.i + 1]) %>% arrange(age_minbin)

      temp1 <- data.frame(state = state.i, sex = sex.i, era = paste0(years[year.i], "-",  years[year.i + 1]),
                          contribution.to.gap.change(type.of.decomp = "COD", 
                                                     decomp.table1 = table1, 
                                                     decomp.table2 = table2, 
                                                     type.of.data = "results"))

      temp2 <- data.frame(state = state.i, sex = sex.i, era = paste0(years[year.i], "-",  years[year.i + 1]),
                          contribution.to.gap.change(type.of.decomp = "Age", 
                                                     decomp.table1 = table3, 
                                                     decomp.table2 = table4,
                                                     type.of.data = "results"))

      cod.df <- rbind(cod.df, temp1)
      age.df <- rbind(age.df, temp2)
    }
  }
}

@@

```{r more_figure_prep, echo = F}

cod.df <- make_pretty_decomp_plot(decomp.table = cod.df, strat.var.1 = "state", 
                        strat.var.2 = "sex", strat.var.3 = "era", 
                        sign.var = "narrowed_gap",
                        decomp.var = "Contribution.to.change",
                        decomp.var.prop = "Contrib.to.change.prop", 
                        partition.bar.var = "COD", type.of.data = "results")

age.df <- make_pretty_decomp_plot(decomp.table = age.df, strat.var.1 = "state", 
                        strat.var.2 = "sex", strat.var.3 = "era", 
                        sign.var = "narrowed_gap",
                        decomp.var = "Contribution.to.change",
                        decomp.var.prop = "Contrib.to.change.prop", 
                        partition.bar.var = "Ages", type.of.data = "results")

@@

```{r, echo = F}
sum1 <- BlackWhite_results %>% filter(year == 1969 & sex == "Male") %>% 
  arrange(LE_wbgap) %>% 
  select(state, LE_wbgap_lcl, LE_wbgap, LE_wbgap_ucl)

@@

```{r, echo =F}
sum2 <- BlackWhite_results %>% filter(year == 1983 & sex == "Male") %>% 
  arrange(LE_wbgap) %>% 
  select(state, LE_wbgap_lcl, LE_wbgap, LE_wbgap_ucl)

@@

```{r, echo =F}
sum3 <- BlackWhite_results %>% filter(year == 1993 & sex == "Male") %>% 
  arrange(LE_wbgap) %>% 
  select(state, LE_wbgap_lcl, LE_wbgap, LE_wbgap_ucl)

sum4 <- BlackWhite_results %>% filter(year == 2013 & sex == "Male") %>% 
  arrange(LE_wbgap) %>% 
  select(state, LE_wbgap_lcl, LE_wbgap, LE_wbgap_ucl)
```

@@

males.era1 <- ggplotly(
  ggplot(subset(cod.df.males, era == "1969-1983"), aes(y = state.gdiff.num.1, x = new.start.1)) +
    geom_rect(aes(xmin = new.start.1, ymin = state.gdiff.num.1 - 0.45, 
                  ymax = state.gdiff.num.1 + 0.45, 
                  xmax = new.finish.1, fill = COD), 
              color = "white") +
    geom_segment(data = subset(cod.df.males, COD == "Cancers" & era == "1969-1983"), aes(x = second.gap, xend = second.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5)) + 
        geom_segment(data = subset(cod.df.males, COD == "Cancers" & era == "1969-1983"), aes(x = first.gap, xend = first.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5), lty = 3) + 
    theme_minimal() + 
    scale_x_continuous(breaks = seq(2, 16, by = 2), labels = seq(2, 16, by = 2)) +
    scale_y_continuous(breaks = 1:y.num.breaks, labels = y.labels)) %>%
  layout(xaxis = list(title = "Life expectancy gap (years): 1969-1983"), yaxis = list(title = NA, autorange = "reversed"))

>>>>

```{r read_data, echo=F}
age_cod_results <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_cod_results.csv")
age_decomp_results <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/age_decomp_results.csv")
cod_decomp_results <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/cod_decomp_results.csv")
cod_change_results <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/cod_change_results.csv")
BlackWhite_results <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/BlackWhite_results.csv")

age_cod_results$COD <- factor(age_cod_results$COD, levels(age_cod_results$COD)[c(3, 2, 4, 5, 6, 1)])
age_decomp_results$age <- factor(age_decomp_results$age, levels = levels(age_decomp_results$age)[c(1, 2, 11, 3:10, 12:19)])
age_cod_results$age <- factor(age_cod_results$age, levels = levels(age_cod_results$age)[c(1, 2, 11, 3:10, 12:19)])

cod_change_results$Cause.of.death <- factor(cod_change_results$Cause.of.death, levels(cod_change_results$Cause.of.death)[c(3, 2, 4, 5, 6, 1)])
```

@@

bw_dat <- BlackWhite_results %>% 
  filter(year %in% c(year1, year2, year3, year4), is.na(LE_wbgap_med) == F) %>%
  select(state, year, LE_wbgap_med, sex) %>%
  arrange(year)

bw_dat <- tidyr::spread(bw_dat, year, LE_wbgap_med) 
names(bw_dat)[3:6] <- c("first.gap","second.gap", "third.gap", "fourth.gap")  
bw_dat.males <- bw_dat %>% filter(sex == "Male") %>%
  mutate(era1.diff = first.gap - second.gap,
         era2.diff = second.gap - third.gap,
         era3.diff = third.gap - fourth.gap,
         state.gdiff.order.1 = reorder(state, first.gap),
         state.gdiff.num.1 = as.numeric(state.gdiff.order.1),
         state.gdiff.order.2 = reorder(state, second.gap),
         state.gdiff.num.2 = as.numeric(state.gdiff.order.2),
         state.gdiff.order.3 = reorder(state, third.gap),
         state.gdiff.num.3 = as.numeric(state.gdiff.order.3)
  )

@@

# cod.df <- data.frame(state = factor(), sex = factor(), era = character(), 
#                      COD = factor(), Contribution.to.change = numeric(), 
#                      Contrib.to.change.prop = numeric(), narrowed_gap = logical())
# 
# age.df <- data.frame(state = factor(), sex = factor(), era = character(), 
#                      age_minbin = numeric(), Contribution.to.change = numeric(), 
#                      Contrib.to.change.prop = numeric(), narrowed_gap = logical())
# 
# years <- c(1969, 1983, 1993, 2013)
# 
# for (state.i in levels(BlackWhite_results$state)) {
#   for (sex.i in c("Male", "Female")) {
#     for (year.i in 1:3) {
#       
#       table1 <- cod_decomp_results %>% filter(state == state.i & sex == sex.i & year == years[year.i]) %>% arrange(COD)
#       table2 <- cod_decomp_results %>% filter(state == state.i & sex == sex.i & year == years[year.i + 1]) %>% arrange(COD)
#       
#       table3 <- age_decomp_results %>% filter(state == state.i & sex == sex.i & year == years[year.i]) %>% arrange(age_minbin)
#       table4 <- age_decomp_results %>% filter(state == state.i & sex == sex.i & year == years[year.i + 1]) %>% arrange(age_minbin)
#       
#       temp1 <- data.frame(state = state.i, sex = sex.i, era = paste0(years[year.i], "-",  years[year.i + 1]),
#                           contribution.to.gap.change(type.of.decomp = "COD", 
#                                                      decomp.table1 = table1, 
#                                                      decomp.table2 = table2, 
#                                                      type.of.data = "results"))
#       
#       temp2 <- data.frame(state = state.i, sex = sex.i, era = paste0(years[year.i], "-",  years[year.i + 1]),
#                           contribution.to.gap.change(type.of.decomp = "Age", 
#                                                      decomp.table1 = table3, 
#                                                      decomp.table2 = table4,
#                                                      type.of.data = "results"))
#       
#       cod.df <- rbind(cod.df, temp1)
#       age.df <- rbind(age.df, temp2)
#     }
#   }
# }

@@

```{r more_figure_prep, echo = F}
cod.df <- cod_change_results %>% select(state, sex, Cause.of.death, Cx_era1_diff_mean, Cx_era2_diff_mean, Cx_era3_diff_mean)
cod.df <- tidyr::gather(cod.df, Cx_era1_diff_mean, Cx_era2_diff_mean, Cx_era3_diff_mean, key = "era", value = "Cx_diff_mean")
ncols <- str_split_fixed(as.character(cod.df$era), "_", 4)
cod.df$era <- as.factor(ncols[ , 2])

cod.df2 <- cod_change_results %>% select(state, sex, Cause.of.death, Cx_p_era1_diff_mean, Cx_p_era2_diff_mean, Cx_p_era3_diff_mean)
cod.df2 <- tidyr::gather(cod.df2, Cx_p_era1_diff_mean, Cx_p_era2_diff_mean, Cx_p_era3_diff_mean, key = "era", value = "Cx_p_diff_mean")
ncols <- str_split_fixed(as.character(cod.df2$era), "_", 5)
cod.df2$era <- as.factor(ncols[ , 3])

cod.df <- merge(cod.df, cod.df2, by = c("state", "sex", "Cause.of.death", "era"))

cod.df$narrowed_gap <- T
cod.df$narrowed_gap[cod.df$Cx_diff_mean < 0 ] <- F

cod.df <- make_pretty_decomp_plot(decomp.table = cod.df, strat.var.1 = "state",
                                  strat.var.2 = "sex", strat.var.3 = "era", sign.var = "narrowed_gap",
                                  decomp.var = "Cx_diff_mean",
                                  decomp.var.prop = "Cx_p_diff_mean",
                                  partition.bar.var = "Cause.of.death",
                                  type.of.data = "results")

@@

```{r, echo = F}
sum1 <- BlackWhite_results %>% filter(year == 1969 & sex == "Male") %>% 
  arrange(LE_wbgap_mean) %>% 
  select(state, LE_wbgap_lcl, LE_wbgap_mean, LE_wbgap_ucl)

@@

```{r, echo =F}
sum2 <- BlackWhite_results %>% filter(year == 1983 & sex == "Male") %>% 
  arrange(LE_wbgap_mean) %>% 
  select(state, LE_wbgap_lcl, LE_wbgap_mean, LE_wbgap_ucl)

@@

```{r, echo =F}
sum3 <- BlackWhite_results %>% filter(year == 1993 & sex == "Male") %>% 
  arrange(LE_wbgap_mean) %>% 
  select(state, LE_wbgap_lcl, LE_wbgap_mean, LE_wbgap_ucl)

sum4 <- BlackWhite_results %>% filter(year == 2013 & sex == "Male") %>% 
  arrange(LE_wbgap_mean) %>% 
  select(state, LE_wbgap_lcl, LE_wbgap_mean, LE_wbgap_ucl)
```

@@

males.era1 <- ggplotly(
  ggplot(subset(cod.df.males, era == "era1"), aes(y = state.gdiff.num.1, x = new.start.1)) +
    geom_rect(aes(xmin = new.start.1, ymin = state.gdiff.num.1 - 0.45, 
                  ymax = state.gdiff.num.1 + 0.45, 
                  xmax = new.finish.1, fill = Cause.of.death), 
              color = "white") +
    geom_segment(data = subset(cod.df.males, Cause.of.death == "Cancers" & era == "era1"), aes(x = second.gap, xend = second.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5)) + 
        geom_segment(data = subset(cod.df.males, Cause.of.death == "Cancers" & era == "era1"), aes(x = first.gap, xend = first.gap, y = state.gdiff.num.1 - 0.5, yend = state.gdiff.num.1 + 0.5), lty = 3) + 
    theme_minimal() + 
    scale_x_continuous(breaks = seq(2, 16, by = 2), labels = seq(2, 16, by = 2)) +
    scale_y_continuous(breaks = 1:y.num.breaks, labels = y.labels)) %>%
  layout(xaxis = list(title = "Life expectancy gap (years): 1969-1983"), yaxis = list(title = NA, autorange = "reversed"))
